---
title: "Implicit Analysis"
output: html_document
params:
  base_dir: ~/Research/Projects/scala-implicits-analysis/corpora/top100
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(fs)
library(tidyverse)
library(ggplot2)
library(DT)

theme_set(theme_minimal())
```

```{r constants, include=FALSE}
IMPLICIT_DECLARATIONS <- path(params$base_dir, "implicit-declarations.csv")
IMPLICIT_CALLSITES <- path(params$base_dir, "implicit-callsites.csv")
IMPLICITS_EXCEPTIONS <- path(params$base_dir, "implicits-exceptions.csv")
CORPUS <- path(params$base_dir, "corpus.csv")
```

```{r aux functions}
is_outlier_min <- function(x) quantile(x, 0.25) - 1.5 * IQR(x)
is_outlier_max <- function(x) quantile(x, 0.75) + 1.5 * IQR(x)
is_outlier <- function(x) {
  (x < is_outlier_min(x)) | (x > is_outlier_max(x))
}
```

```{r load data, include=FALSE}
declarations <- read_csv(IMPLICIT_DECLARATIONS)
callsites <- read_csv(IMPLICIT_CALLSITES)
corpus <- read_csv(CORPUS)
```

## Checking data consistency

### Declarations

```{r}
count(declarations, project_id) %>% 
  rename(bin=n) %>%
  left_join(select(corpus, project_id, stats=implicit_declarations), by="project_id") %>% 
  filter(bin-stats != 0)
```

### Call sites

```{r}
count(callsites, project_id) %>% 
  rename(bin=n) %>%
  left_join(select(corpus, project_id, stats=implicit_callsites), by="project_id") %>% 
  filter(bin-stats != 0)
```


```{r}
stopifnot(nrow(callsites) == sum(corpus$implicit_callsites, na.rm = T))
```


```{r library classification}
TEST_LIBS_GROUP_ID <-
  c(
    "org.scalacheck",
    "org.scalamock",
    "org.scalatest",
    "org.specs2"
  )

# TODO:    "com.lihaoyi" %% "utest"

is_test_lib <- function(group_id) group_id %in% TEST_LIBS_GROUP_ID 
is_scala_stdlib <- function(group_id) startsWith(group_id, "org.scala-lang")
```

```{r path classificaton}
```


```{r only use completed projects}
projects <- filter(corpus, implicits_exit_code == 0)
callsites <- semi_join(callsites, projects, by="project_id")
declarations <- semi_join(declarations, projects, by="project_id")
```

```{r stats functions}
apps <- filter(projects, scaladex==FALSE)
libs <- filter(projects, scaladex==TRUE)

add_stat <- function(key, expr) {
  expr <- substitute(expr)
  fun <- function(x) eval(expr, envir=list2env(list(.=x)))
  tibble(key, apps=fun(apps), libs=fun(libs), all=fun(projects))
}

add_stat_num <- function(key, expr) {
  expr <- substitute(expr)
  fun <- function(x) eval(expr, envir=list2env(list(.=x)))
  
  stats <- list(sum, mean, median, sd)
  names(stats) <- c(
    str_glue("Number of {key}"), 
    str_glue("Average of {key} per project"), 
    str_glue("Median of {key}"), 
    str_glue("SD of {key}")
  )
  
  stats %>% imap_dfr(~tibble(key=.y, apps=.x(fun(apps)), libs=.x(fun(libs)), all=.x(fun(projects))))
}

make_stats <- function(...) {
  all <- list(...)
  bind_rows(all)
}
```

## Summary

```{r summary}
stats <- make_stats(
  add_stat("Number of projects", nrow(.)),
  add_stat("Number of projects declaring implicits", nrow(filter(., implicit_declarations > 0))),
  add_stat("Number of projects not declaring implicits", nrow(filter(., implicit_declarations == 0))),
  add_stat("Number of projects using implicits", nrow(filter(., implicit_callsites > 0))),
  add_stat("Number of projects not using implicits", nrow(filter(., implicit_callsites == 0))),
  add_stat("Number of projects failing implicit extraction", nrow(filter(., !is.na(implicit_failure)))),
  add_stat_num("call sites", .$callsites),
  add_stat_num("problems extracting exceptions", .$implicit_problems),
  add_stat_num("implicit call sites", .$implicit_callsites),
  add_stat_num("implicit declarations", .$implicit_declarations),
  add_stat_num("github stars", .$github_stars)
)

stats %>% datatable(options=list(paging=FALSE, searching=FALSE, info=FALSE))
```

## Problems with implicit extraction

```{r}
exceptions <- read_csv(IMPLICITS_EXCEPTIONS, col_types="ccccc")
```

```{r}
classified_exceptions <- 
  mutate(
    exceptions,
    class=case_when(
      cause == "MissingSymbolException" & str_detect(message, "symbol: local\\d+") ~ "missing-local-symbol",
      cause == "MissingSymbolException" & str_detect(message, "at Range\\(\\d+,\\d+,\\d+,\\d+\\)") ~ "missing-symbol-at-range",
      cause == "MissingSymbolException" ~ "missing-symbol-other",
      cause == "LoadingMetadataException" & str_detect(message, "No module found") & str_detect(message, "multi-jvm") ~ "missing-module-multijvm",
      cause == "LoadingMetadataException" & str_detect(message, "No module found") ~ "missing-module",
      cause == "LoadingMetadataException" ~ "metadata-loading",
      cause == "NotImplementedError" ~ "not-implemented",
      cause == "UnexpectedElementException" ~ "unexpected-element",
      cause == "UnsupportedElementException" ~ "unsupported-element",
      TRUE ~ "unclassified"
    )
  )
```


### Per project summary

```{r}
count(classified_exceptions, project_id) %>% arrange(desc(n)) %>% datatable()
```

### Per exception summary

```{r}
count(classified_exceptions, cause) %>% arrange(desc(n)) %>% datatable()
```

### Per class summary

```{r}
count(classified_exceptions, class) %>% arrange(desc(n)) %>% datatable()
```

### Per project and class summary

```{r}
count(classified_exceptions, project_id, class) %>% arrange(desc(n)) %>% datatable()
```

### Missing local symbols

```{r}
filter(classified_exceptions, class=="missing-local-symbol") %>% count(project_id) %>% arrange(desc(n)) %>% datatable()
```

#### Missing symbols at range

```{r}
filter(classified_exceptions, class=="missing-symbol-at-range") %>% count(project_id) %>% arrange(desc(n)) %>% datatable()
```

#### Missing other symbols

```{r}
filter(classified_exceptions, class=="missing-symbol-other") %>% count(project_id) %>% arrange(desc(n)) %>% datatable()
```

#### Per project and symbol

```{r}
filter(classified_exceptions, class=="missing-symbol-other") %>% 
  transmute(project_id, symbol=str_replace_all(message, ".*symbol: (.*)$", "\\1")) %>%
  count(project_id, symbol) %>% 
  arrange(desc(n)) %>% 
  datatable()
```

### Unclassified

```{r}
uncategorized_exceptions <- filter(classified_exceptions, class=="unclassified")
```


## Source code 

- add the outlier flag, color the dot and list the projects

```{r}
scatter_plot_sloc <- function(var, y) {
  data <- projects %>%
  mutate(
    r=round(!!var/metadata_scala_code, 2),
    outlier=is_outlier(r),
    label=if_else(outlier, str_glue("{project_id} ({r})"), as.character(NA))
  )
  
  ggplot(data, aes(x=metadata_scala_code, y=!!var, label=label)) + 
  geom_point(aes(color=factor(outlier))) +
  geom_text(size=3, check_overlap = T, vjust=1.5, na.rm = TRUE) +
  scale_x_log10(breaks=c(1,10,100,1000,10000,100000), labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) + 
  scale_colour_manual(values = c("TRUE"="red", "FALSE"="black"), labels=c("TRUE"="Yes", "FALSE"="No"), guide=FALSE) +
  labs(
    x="Source lines of code (log)", 
    y=y, 
    color="Outlier", 
    subtitle=str_glue("Red indicates outlier of y/x, outside of ({is_outlier_min(data$r)}, {is_outlier_max(data$r)})")
  )
}
```

```{r}
scatter_plot_sloc(quo(implicit_callsites), "Implicit call sites")
```

```{r}
scatter_plot_sloc(quo(implicit_declarations), "Implicit declarations")
```

```{r}
scatter_plot_sloc(quo(callsites), "Call sites")
```

## Implicit conversions
